Appendix to 'cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data

### 2022-06-26 jss/appendix.Rmd
script on appendix

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(cubble)
library(patchwork)
```

@@

```{r echo = TRUE}
cluster_nested <- station_nested %>% switch_key(cluster)
```

@@

```{r eval = FALSE, echo = FALSE}
coords <- cbind(prcp_aus$long, prcp_aus$lat)
dist_raw <- geosphere::distm(coords, coords)

@@

Figure 9 in this section can also be made using `cubble` and `leaflet`, in which case the temperature range is displayed as a small subplot upon clicking on the map. This procedure involves first creating the popup plots from the long form cubble as a vector and then adding these plots to a leaflet map created from the nested cubble, with `leafpop::addPopupGraphs()`:

@@

Figure \ref{fig:interactive-popup} shows the same information as Figure 9 but using leaflet and popups.

```{r interactive-popup, fig.align='center', echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map.", fig.show='hold'}
knitr::include_graphics(here::here("figures/popup-mid.png"))
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(cubble)
library(patchwork)
library(leaflet)
library(leafpop)
```

@@

```{r eval = FALSE, echo = TRUE}
cluster_nested <- station_nested %>% switch_key(cluster)
```

@@

```{r eval = FALSE, echo = FALSE}
# The kmeans clustering in station_nested data can takes some time to run, 
# This chunk runs the algorithm and save the result as staiton_nested for further use
coords <- cbind(prcp_aus$long, prcp_aus$lat)
dist_raw <- geosphere::distm(coords, coords)

@@

Figure \ref{fig:interactive-popup} in this section can also be made using `cubble` and `leaflet`, in which case the temperature range is displayed as a small subplot upon clicking on the map. This procedure involves first creating the popup plots from the long form cubble as a vector and then adding these plots to a leaflet map created from the nested cubble, with `leafpop::addPopupGraphs()`:

@@

Figure \ref{fig:interactive-popup} shows the same information as Figure 9 but using leaflet and popups [@leafpop].

```{r eval = FALSE, echo = FALSE}
clean <-  climate_full %>%
  filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
                    "ASN00097053", "ASN00040856", "ASN00015548")) %>%
  face_temporal() %>%
  mutate(month = lubridate::month(date)) %>% 
  group_by(month) %>% 
  summarise(tmax = mean(tmax, na.rm = TRUE),
            tmin = mean(tmin, na.rm = TRUE),
            diff = mean(tmax - tmin, na.rm = TRUE)) %>% 
  ungroup(month) %>% 
  face_spatial() %>%
  mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE)) %>% 
  face_temporal() %>% 
  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>% 
  unfold(name)

df_id <- clean$id %>% unique()
p <- purrr::map(1:length(df_id), function(i){
  dt <- clean %>% filter(id == df_id[i])
  dt %>% 
    ggplot(aes(x = dummy_date, ymin = tmin, ymax = tmax, group = id)) + 
    geom_ribbon(color = "grey50", fill = "grey50", alpha = 0.5) + 
    theme_bw() + 
    scale_x_date(date_labels = "%b", date_breaks = "1 month") + 
    labs(x = "Month", y = "Temperature") + 
    ylim(-10, 50) + 
    ggtitle(paste0(stringr::str_to_title(unique(dt$name))))
})

nested <- face_spatial(clean) 
domain <- nested$temp_diff_var
pal <- colorNumeric(
  colorspace::sequential_hcl(
    "Rocket",  n = 7, cmax = 90, rev = TRUE, 
    c2 = 40, l2= 85, c1 = 20, l1 = 30),
  domain = domain)

leaflet(nested) %>% 
  addTiles() %>% 
  addCircleMarkers(color = ~pal(domain), radius = 0.1, 
                   fillOpacity = 1, opacity = 1, 
                   popup = popupGraph(p, width = 300, height =200)) 
```

```{r interactive-popup, fig.align='center', echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map.", fig.show='hold'}
knitr::include_graphics(here::here("figures/popup.png"))
```
